Demo arXiv template

### 2023-07-04 tidyindex.qmd
outline the challenges with evaluating indexes and our proposal

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(width = 90)
library(tidyverse)
library(latex2exp)
library(patchwork)
library(tidyindex)
library(lmomco)
library(lubridate)
library(SPEI)
library(tsibble)
library(GGally)
library(patchwork)
```

@@

<!-- FYI:  -->

<!--   - A measure is a variable, but is not an index.  -->
<!--   - Not all multivariate quantities are indexes (i.e. those remote sensor measures). It is possible that an index is constructed from a single variable (i.e. SPI). -->
<!--   - indexes is not a single/ set of measures -->

@@

*see if there is any paper describing this type of pains*

*useful to reference tidy data and tidy model that makes the workflow on modelling tidy somewhere in introduction*

Currently, index construction lacks a standardised workflow. 
It is often up to researchers or research institutions to decide whether to provide open source code on the new indexes, what would be the best user interface for other researchers to use the new indexes, and how easily the new indexes can be compared with other existing indexes. 
This makes the computation lack transparency and indexes cumbersome to experiment with:

  - Researchers who wish to validate the indexes calculated from large institutes need to reinvent the wheels themselves since the source code used for computing is often not available for public consumption;
  - Open-source code provided by research groups has a narrow margin for exploring other options outside the provided;
  - Similar steps used by different indexes are difficult to spot since the design of the user interface for indexes often includes all the steps under a single function call; and
  - It is generally hard to inspect intermediate results during the index construction if users wish to check the output of a certain step.

<!-- Institutes like National Oceanic and Atmospheric Administration (NOAA) and United States Drought Monitor (USDM) moderate indexes in their domains. -->
<!-- They provide public information to describe the methodology used in the calculation, publish the index values for public consumption while the source code used for computing is often not available for public consumption.  -->

<!-- For example, many indexes involve a step to fit variables into a particular distribution. -->
<!-- Index researchers may provide options for different fits, but it is generally hard to experiment with options outside the provided distributions. -->

<!-- For example, many indexes needs standardisation in its workflow, it can be used to rescale the input variables or the index series. -->
<!-- But the current construction does not make it clear through its user interface but sweep it, along with all the other steps in the index construction, under a single function call.  -->

@@

This paper proposes a data pipeline for index construction. 
By recognising the common steps shared by many indexes, we develop a pipeline that breaks down index construction into multiple modules and allow operations in various modules to be combined like building blocks to construct indexes.
The pipeline approach is general while adaptable to most index construction. 
It allows indexes to be created, studied, and compared in a structured tidy form and enables statistical analysis of indexes to be performed easily: 
More specifically, it enables researchers to 1) validate the indexes calculated from external organisations, 2) unify various indexes under the same framework for computing, 3) swap or adjust individual steps in the index construction to study their contribution, 4) calculate uncertainty on indexes through bootstrap or others, 5) enhance existing indexes through comparing and studying their statistical properties, and finally, 6) propose new indexes from combining different steps in existing indexes.

@@

This work is of interest to researchers actively developing new indexes since it encourages new indexes to be delivered in an easy-to-reproduce design. 
It would also provide analysts who wish to compute a range of indexes in their analysis a uniform interface to build relevant indexes from raw data. 
For statisticians and software developing engineers, this work frames the process of index construction in a more user-oriented workflow and could motivate similar research for other process in scientific computing.

The rest of the paper is structured as follows: @sec-data-pipeline-in-r reviews the concept of data pipeline in R. 
The pipeline framework for index construction is presented in @sec-a-pipeline-for-building-statistical-indexes.  @sec-incorporating-new-buliding-blocks-into-the-pipeline explains how to include a new building block in each pipeline module. 
Examples are given in @sec-examples to demonstrate the index construction with the pipeline built.

# Data pipeline {#sec-data-pipeline-in-r}

@@

# A pipeline for building statistical indexes {#sec-a-pipeline-for-building-statistical-indexes}

@@

```{r}
knitr::include_graphics(here::here("figures/toy-example-spei.png"))
```

@@

# Incorporating alternative methods into the pipeline components {#sec-incorporating-new-buliding-blocks-into-the-pipeline}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(width = 70)
library(tidyverse)
library(latex2exp)
library(patchwork)
library(tidyindex)
library(lmomco)
library(lubridate)
library(SPEI)
library(tsibble)
library(GGally)
library(patchwork)
```

@@

@@

  - Difficult to reach consensus on composite/multivariate indexes. 
  - Given the complexity of multivariate indexes as compared to univariate indexes, it raises more doubt to persuade people to use it. 
  - Typical ways to evaluate new indexes is to compare it with the standard indexes (i.e. SPI in drought index) or a response variable (i.e. crop yield for monitoring agricultural drought). 
  - However, more doubt can still raises: Whether the variable used to construct the index is a good set of variables. 
  - Does the index series capture the variables measures. Whether other ways of combining the variables results in difference in the index. 
  - Being able to answer these questions help to prove the robustness, extendibility and dimensionality of the index, as proposed by xxx as criteria for (selecting?) indexes.

@@

  - In order to evaluate indexes, indexes need to be broken down into pieces that it constitutes and then we can analyse how each component will affect the index. 
  - The OECD handbook [@oecd_handbook_2008] provides technical guidelines to construct composite indexes for comparing country performance for policy-making. 
  This paper identifies a more general pipeline framework where statistical indexes (of all kinds) can be built from. 
  Based on the pipeline, an R package is built to validate the indexes, [more details on the functionality here] from understanding multivariate relationships to uncertainty analysis.

<!-- The pipeline approach is general while adaptable to most index construction.  -->
<!-- It allows indexes to be created, studied, and compared in a structured tidy form and enables statistical analysis of indexes to be performed easily:  -->

@@

  - This work provides researchers actively developing new indexes with tools to evaluate the proposed indexes. 
  - Helping index analysts to identify weakness in existing indexes for methodology improvement.

The rest of the paper is structured as follows: @sec-pipeline reviews the concept of data pipeline in R. The pipeline framework for index construction is presented in @sec-index-pipeline. @sec-dev explains how to include a new building block in each pipeline module. Examples are given in @sec-examples to demonstrate the index construction with the pipeline built.

# Data pipeline {#sec-pipeline}

@@

# A pipeline for building statistical indexes {#sec-index-pipeline}

@@

```{r width = "100%"}
knitr::include_graphics(here::here("figures/toy-example-spei.png"))
```

@@

# Incorporating alternative methods into the pipeline components {#sec-dev}

### 2023-07-04 tidyindex.tex
outline the challenges with evaluating indexes and our proposal

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne, VIC, Australia
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[borderline west={3pt}{0pt}{shadecolor}, sharp corners, interior hidden, frame hidden, enhanced, breakable, boxrule=0pt]}{\end{tcolorbox}}\fi

@@

\emph{see if there is any paper describing this type of pains}

\emph{useful to reference tidy data and tidy model that makes the
workflow on modelling tidy somewhere in introduction}

Currently, index construction lacks a standardised workflow. It is often
up to researchers or research institutions to decide whether to provide
open source code on the new indexes, what would be the best user
interface for other researchers to use the new indexes, and how easily
the new indexes can be compared with other existing indexes. This makes
the computation lack transparency and indexes cumbersome to experiment
with:

\begin{itemize}
\tightlist
\item
  Researchers who wish to validate the indexes calculated from large
  institutes need to reinvent the wheels themselves since the source
  code used for computing is often not available for public consumption;
\item
  Open-source code provided by research groups has a narrow margin for
  exploring other options outside the provided;
\item
  Similar steps used by different indexes are difficult to spot since
  the design of the user interface for indexes often includes all the
  steps under a single function call; and
\item
  It is generally hard to inspect intermediate results during the index
  construction if users wish to check the output of a certain step.
\end{itemize}

@@

This paper proposes a data pipeline for index construction. By
recognising the common steps shared by many indexes, we develop a
pipeline that breaks down index construction into multiple modules and
allow operations in various modules to be combined like building blocks
to construct indexes. The pipeline approach is general while adaptable
to most index construction. It allows indexes to be created, studied,
and compared in a structured tidy form and enables statistical analysis
of indexes to be performed easily: More specifically, it enables
researchers to 1) validate the indexes calculated from external
organisations, 2) unify various indexes under the same framework for
computing, 3) swap or adjust individual steps in the index construction
to study their contribution, 4) calculate uncertainty on indexes through
bootstrap or others, 5) enhance existing indexes through comparing and
studying their statistical properties, and finally, 6) propose new
indexes from combining different steps in existing indexes.

@@

This work is of interest to researchers actively developing new indexes
since it encourages new indexes to be delivered in an easy-to-reproduce
design. It would also provide analysts who wish to compute a range of
indexes in their analysis a uniform interface to build relevant indexes
from raw data. For statisticians and software developing engineers, this
work frames the process of index construction in a more user-oriented
workflow and could motivate similar research for other process in
scientific computing.

The rest of the paper is structured as follows:
Section~\ref{sec-data-pipeline-in-r} reviews the concept of data
pipeline in R. The pipeline framework for index construction is
presented in
Section~\ref{sec-a-pipeline-for-building-statistical-indexes}.
Section~\ref{sec-incorporating-new-buliding-blocks-into-the-pipeline}
explains how to include a new building block in each pipeline module.
Examples are given in Section~\ref{sec-examples} to demonstrate the
index construction with the pipeline built.

\hypertarget{sec-data-pipeline-in-r}{%
\section{Data pipeline}\label{sec-data-pipeline-in-r}}

@@

\hypertarget{sec-a-pipeline-for-building-statistical-indexes}{%
\section{A pipeline for building statistical
indexes}\label{sec-a-pipeline-for-building-statistical-indexes}}

@@

\hypertarget{sec-incorporating-new-buliding-blocks-into-the-pipeline}{%
\section{Incorporating alternative methods into the pipeline
components}\label{sec-incorporating-new-buliding-blocks-into-the-pipeline}}

@@

Data: 
# A tibble: 191 x 13
      id country  life_exp exp_sch avg_sch gni_pc   sch index0 index1 index2 index3 index4
   <dbl> <chr>       <dbl>   <dbl>   <dbl>  <dbl> <dbl>  <dbl>  <dbl>  <dbl>  <dbl>  <dbl>
 1     1 Switzer~    0.984   0.917   0.924  0.983 0.920  0.962  0.963   1.67  0.978  0.933
 2     2 Norway      0.973   1       0.867  0.978 0.933  0.961  0.961   1.66  0.969  0.942
 3     3 Iceland     0.964   1       0.918  0.955 0.959  0.959  0.959   1.66  0.963  0.959
 4     4 Hong Ko~    1       0.960   0.815  0.973 0.887  0.952  0.953   1.65  0.986  0.907
 5     5 Austral~    0.993   1       0.848  0.936 0.924  0.951  0.951   1.65  0.980  0.932
 6     6 Denmark     0.944   1       0.864  0.967 0.932  0.948  0.948   1.64  0.945  0.937
 7     7 Sweden      0.969   1       0.841  0.952 0.920  0.947  0.947   1.64  0.962  0.928
 8     8 Ireland     0.954   1       0.772  1     0.886  0.945  0.947   1.64  0.952  0.904
 9     9 Germany     0.933   0.945   0.939  0.952 0.942  0.942  0.942   1.63  0.936  0.942
10    10 Netherl~    0.949   1       0.839  0.956 0.919  0.941  0.941   1.63  0.947  0.926
# i 181 more rows
# i 1 more variable: index5 <dbl>
\end{verbatim}

@@

>>>>

\thanks{CONTACT: H. Sherry
Zhang. Email: \href{mailto:huize.zhang@monash.edu}{\nolinkurl{huize.zhang@monash.edu}}. }
\begin{document}
\captionsetup{labelsep=space}
\maketitle
\textsuperscript{1} Department of Econometrics and Business
Statistics, Monash University, Melbourne, VIC, Australia
\begin{abstract}
\begin{itemize}
\tightlist
\item
  indexes, useful, quantify severity, early monitoring,
\item
  A huge number of indexes have been proposed by domain experts,
  however, a large majority of them are not being adopted, reused, and
  compared in research or in practice.
\item
  One of the reasons for this is the plenty of indexes are quite complex
  and there is no obvious easy-to-use implementation to apply them to
  user's data.
\item
  The paper describes a general pipeline framework to construct indexes
  from spatio-temporal data,
\item
  This allows all the indexes to be constructed through a uniform data
  pipeline and different indexes to vary on the details of each step in
  the data pipeline and their orders.
\item
  The pipeline proposed aim to smooth the workflow of index construction
  through breaking down the complicated steps proposed by various
  indexes into small building blocks shared by most of the indexes.
\item
  The framework will be demonstrated with drought indexes as examples,
  but appliable in general to environmental indexes constructed from
  multivariate spatio-temporal data
\end{itemize}
\end{abstract}
\begin{keywords}
\def\sep{;\ }
indexes\sep data pipeline\sep 
software design
\end{keywords}
\ifdefined\Shaded\renewenvironment{Shaded}{\begin{tcolorbox}[boxrule=0pt, frame hidden, sharp corners, borderline west={3pt}{0pt}{shadecolor}, interior hidden, breakable, enhanced]}{\end{tcolorbox}}\fi

@@

\begin{itemize}
\tightlist
\item
  Difficult to reach consensus on composite/multivariate indexes.
\item
  Given the complexity of multivariate indexes as compared to univariate
  indexes, it raises more doubt to persuade people to use it.
\item
  Typical ways to evaluate new indexes is to compare it with the
  standard indexes (i.e.~SPI in drought index) or a response variable
  (i.e.~crop yield for monitoring agricultural drought).
\item
  However, more doubt can still raises: Whether the variable used to
  construct the index is a good set of variables.
\item
  Does the index series capture the variables measures. Whether other
  ways of combining the variables results in difference in the index.
\item
  Being able to answer these questions help to prove the robustness,
  extendibility and dimensionality of the index, as proposed by xxx as
  criteria for (selecting?) indexes.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  In order to evaluate indexes, indexes need to be broken down into
  pieces that it constitutes and then we can analyse how each component
  will affect the index.
\item
  The OECD handbook (OECD, European Union, and Joint Research Centre -
  European Commission 2008) provides technical guidelines to construct
  composite indexes for comparing country performance for policy-making.
  This paper identifies a more general pipeline framework where
  statistical indexes (of all kinds) can be built from. Based on the
  pipeline, an R package is built to validate the indexes, {[}more
  details on the functionality here{]} from understanding multivariate
  relationships to uncertainty analysis.
\end{itemize}

@@

\begin{itemize}
\tightlist
\item
  This work provides researchers actively developing new indexes with
  tools to evaluate the proposed indexes.
\item
  Helping index analysts to identify weakness in existing indexes for
  methodology improvement.
\end{itemize}

The rest of the paper is structured as follows:
Section~\ref{sec-pipeline} reviews the concept of data pipeline in R.
The pipeline framework for index construction is presented in
Section~\ref{sec-index-pipeline}. Section~\ref{sec-dev} explains how to
include a new building block in each pipeline module. Examples are given
in Section~\ref{sec-examples} to demonstrate the index construction with
the pipeline built.

\hypertarget{sec-pipeline}{%
\section{Data pipeline}\label{sec-pipeline}}

@@

\hypertarget{sec-index-pipeline}{%
\section{A pipeline for building statistical
indexes}\label{sec-index-pipeline}}

@@

\hypertarget{sec-dev}{%
\section{Incorporating alternative methods into the pipeline
components}\label{sec-dev}}

@@

Data: 
# A tibble: 191 x 13
      id country   life_exp exp_sch avg_sch gni_pc   sch index0 index1
   <dbl> <chr>        <dbl>   <dbl>   <dbl>  <dbl> <dbl>  <dbl>  <dbl>
 1     1 Switzerl~    0.984   0.917   0.924  0.983 0.920  0.962  0.963
 2     2 Norway       0.973   1       0.867  0.978 0.933  0.961  0.961
 3     3 Iceland      0.964   1       0.918  0.955 0.959  0.959  0.959
 4     4 Hong Kon~    1       0.960   0.815  0.973 0.887  0.952  0.953
 5     5 Australia    0.993   1       0.848  0.936 0.924  0.951  0.951
 6     6 Denmark      0.944   1       0.864  0.967 0.932  0.948  0.948
 7     7 Sweden       0.969   1       0.841  0.952 0.920  0.947  0.947
 8     8 Ireland      0.954   1       0.772  1     0.886  0.945  0.947
 9     9 Germany      0.933   0.945   0.939  0.952 0.942  0.942  0.942
10    10 Netherla~    0.949   1       0.839  0.956 0.919  0.941  0.941
# i 181 more rows
# i 4 more variables: index2 <dbl>, index3 <dbl>, index4 <dbl>,
#   index5 <dbl>
\end{verbatim}

@@

\leavevmode\vadjust pre{\hypertarget{ref-oecd_handbook_2008}{}}%
OECD, European Union, and Joint Research Centre - European Commission.
2008. \emph{Handbook on {Constructing} {Composite} {Indicators}:
{Methodology} and {User} {Guide}}. OECD.
\url{https://doi.org/10.1787/9789264043466-en}.
